---
title: "Percentage of Population Living Under 2011 PPP$ 1.90 per Day"
output:
  github_document: default
  html_notebook: default
---


```{r setup}
library(tidyverse)
```


# Percentage of Population Living Under 2011 PPP$ 1.90 per Day
## Data import

Data downloaded from https://apihighways.org/data-sets/6a6de0b8-e544-495a-93ab-e8bc3c59fb20

```{r}
df <- read_csv("./data/poverty-190/csv/poverty-190.csv")
df                                                         # 3900 rows : 30 years x 130 countries

entity <- read_csv("./data/poverty-190/csv/entity.csv")
entity
```


## Data cleaning : Headings

```{r}
# r interprets a variable containing a 'minus sign' as an expression. So (for `df`) need to change column headings `value-1`, `value-2`, `value-3`. And to be able to do that, need to surround them with ticks

df <- df %>% dplyr::rename(entity_name = `entity-name`, value_1 = `value-1`, value_2 = `value-2`, value_3 = `value-3`)

entity <- entity %>% dplyr::rename(long_name = `long-name`, iso_alpha_2 = `iso-alpha-2`, iso_alpha_3 = `iso-alpha-3`, income_group = `income-group`, dac_id = `dac-id`, donor_recipient_type = `donor-recipient-type`, has_domestic_data = `has-domestic-data`)
```


## Basic EDA

```{r}
df %>% count(id)                    # 130 rows
df %>% filter(is.na(value_1))       # 6 rows (Solomon Islands : 1984-1989)

entity %>% count(id)                # 333 rows (assume the 130 countries in `df` and 203 'other')
entity %>% count(id) %>% count(n)   # 1 row (ie. there were no duplications in the 333 rows)
entity %>% count(type)
```

Column names in `df` :

* `id` <chr> : country (130 of them)
* `entity_name` <chr> : country (130 of them)
* `year` <int> : year (30, from 1984-2013)
* `value_1` <dbl> : presumably Percentage of Population (with 10% shown as 10, not 0.1)
* `value_2` <dbl> : same as `value_1`
* `value_3` <dbl> : same as `value_1`, except it has 2212 missing values (instead of 6)

Column names in `entity` :

* `id` <chr> : assume it is the 130 countries in `df` and 203 'other'
* `type` <chr> : country (252); organisation (49); region (18); NA (14) ...= 333
* `long_name` <chr> :
* `region` <chr> :
* `iso_alpha_2` <chr> :
* `iso_alpha_3` <chr> :
* `neighbors` <chr> :
* `income_group` <chr> :
* `slug` <chr> :
* `dac_id` <chr> :
* `donor_recipient_type` <chr> :
* `has_domestic_data` <int> :
* `name` <chr> :



```{r}
df %>%
  ggplot(aes(value_1, value_3)) +
  geom_point()                                    # `value_3` matches `value_1` or is `NA` ...!?

df %>% filter(entity_name == "Uganda") %>%
  ggplot(aes(year, value_1)) +
  geom_line()
```


## Combine the dataframes

```{r}
joined <- df %>%
  left_join(entity, by = "id")

joined                                      # 3900 rows x 18 columns
```



```{r}
joined %>% filter(region == "south-of-sahara", year == 2013) %>%
  ggplot(aes(x = reorder(entity_name, value_1), y = value_1, colour = value_1 >= 50)) +
  geom_point() +
  coord_flip()

# 50 is an arbitrary choice, just to reduce the number of countries. (It can be changed)
```

```{r}
joined %>% filter(region == "south-of-sahara", value_1 >= 50) %>%
  ggplot(aes(x = year, y = value_1)) +
  geom_line() +
  facet_wrap(~ entity_name)
```

The chart above is (somewhat) interesting, but isn't what I intended. It shows the 30 countries (of 46) in `region == "south-of-sahara"` that had `value_1 >= 50` in any of the years (and what their `value_1` was in those years only). What I want to see is the 15 countries from the top of the chart above that one (blue dots), which are all those with `value_1 >= 50` in 2013 (and what their `value_1` was in every year)

...I think I can do that with a filtering join (joining to a list of countries, only) :
```{r}
`south-of-sahara_2013_50%` <- joined %>% 
  filter(region == "south-of-sahara", year == 2013, value_1 >= 50) %>%
  select(entity_name)

joined %>% semi_join(`south-of-sahara_2013_50%`) %>%
  ggplot(aes(x = year, y = value_1)) +
  geom_line() +
  facet_wrap(~ entity_name)
```






